---
title: "Figure S4"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Figure S4}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

```{r include = FALSE}
# switch to the project-specific library if you have one (e.g. in the companion
# docker container)
library(switchr)
if (file.exists(file.path(switchrBaseDir(), "supt4h1"))) {
  switchrNoUnload(TRUE)
  switchTo("supt4h1")
}
```

```{r include=FALSE}
library(SUPT4H1)
library(dplyr)
library(ggplot2)
library(limma)
library(tibble)
theme_set(theme_linedraw(base_size = 16) +
          theme(legend.key = element_blank()))
```

```{r}
kFdr <- 0.1
gene_sets <- reactome_sets()
```

## Figure S4A: SUPT4H1 RNAi effect in HEK293 cells

```{r}
v <- fit_hek293(normalize = "spike-in", output = "v")
fit <- fit_hek293(normalize = "spike-in", output = "fit")
tt <- topTable(fit, coef = "SUPT4H1", n = Inf)
```

```{r}
reactome_res <- camera(v, ids2indices(gene_sets, row.names(v)),
                       contrast = match("SUPT4H1", colnames(v$design))) %>%
  tibble::rownames_to_column("set_name") %>%
  dplyr::arrange(PValue)
```

The following table shows all gene sets with FDR < `r kFdr`.

```{r}
reactome_res  %>%
  dplyr::filter(FDR < kFdr) %>%
  camera_table()
caption <- "SUPT4H1 RNAi effect in HEK293 cells"
```

```{r fig.width=3.5, fig.height=4, fig.show='hold', fig.cap=caption, cache=FALSE}
plot_gsea(tt, gene_sets, "Peptide chain elongation", xlim = c(-2, 2)) + 
  theme(legend.position = "bottom", plot.title = element_text(size = 14))
```

## Figure S4B: SUPT4H1 RNAi effect in fibroblasts

```{r}
v <- fit_fibroblasts(normalize = "spike-in", output = "v")
fit <- fit_fibroblasts(normalize = "spike-in", output = "fit")
tt <- topTable(fit, coef = "SUPT4H1", n = Inf)
```

```{r}
reactome_res <- camera(v, ids2indices(gene_sets, row.names(v)),
                       contrast = match("SUPT4H1", colnames(v$design))) %>%
  tibble::rownames_to_column("set_name") %>%
  dplyr::arrange(PValue)
```

The following table shows all gene sets with FDR < `r kFdr`.

```{r}
reactome_res  %>%
  dplyr::filter(FDR < kFdr) %>%
  camera_table()
caption <- "SUPT4H1 RNAi effect in fibroblasts"
```

```{r fig.width=3.5, fig.height=4, fig.show='hold', fig.cap=caption, cache=FALSE}
plot_gsea(tt, gene_sets, "Peptide chain elongation", xlim = c(-5, 5)) + 
  theme(legend.position = "bottom", plot.title = element_text(size = 14))
```
