---
title: "Figure S2"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Figure S2}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

```{r eval=TRUE, include=FALSE}
# switch to the project-specific library if you have one (e.g. in the companion
# docker container)
library(switchr)
if (file.exists(file.path(switchrBaseDir(), "supt4h1"))) {
  switchrNoUnload(TRUE)
  switchTo("supt4h1")
}
```

```{r include=FALSE}
library(SUPT4H1)
library(dplyr)
library(edgeR)
library(ggplot2)
library(limma)
library(stringr)
library(tidyr)
```

```{r}
kLabels = list(
  SPT4 = "SUPT4H1\nsiRNA", 
  Scramble = "Scramble\nsiRNA",
  Control = "Control cell line", 
  DARP1 = "DARP cell line 1",
  DARP2 = "DARP cell line 2",
  Scramble_Control_mix2_A = "Scramble siRNA, Control cells",
  Scramble_DARP1_mix2_A = "Scramble siRNA, DARP1 cells",
  Scramble_DARP2_mix2_A = "Scramble siRNA, DARP2 cells",
  SPT4_Control_mix1_A = "SUPT4H1 siRNA, Control cells",
  SPT4_DARP1_mix1_A = "SUPT4H1 siRNA, DARP1 cells",
  SPT4_DARP2_mix1_A = "SUPT4H1 siRNA, DARP2 cells")
theme_set(theme_linedraw(16))
```

```{r}
dge <- load_dataset("HEK293")
```

## Figure S2A

Boxplot of estimaged log2 fold-changes for the ERCC spike-ins in the HEK293
cell model after global normalization.

```{r}
fit_global <- fit_hek293(dge, normalize = "global")
tt <- topTable(fit_global, coef = "SUPT4H1", n = Inf)
```

```{r fig.height = 5, fig.width = 5.5, fig.show='hold', fig.cap="Figure S2A: SUPT4H1 effect (Global normalization)", cache=FALSE}
plot_boxes(tt, labels = TRUE, cex = 5, nudge_y = 0.5, show_spt4 = TRUE,
           nudge_spt4 = 0.6, ylim = c(-5, 2)) +
  ggtitle("Global normalization") +
  ylab("Mean observed fold change (log2)") +
  theme(legend.position = "none", plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))
```

## Figure S2B

MA-plot of estimaged log2 fold-changes for the ERCC spike-ins in the HEK293
cell model after global normalization.

```{r fig.height = 4, fig.width = 5.5, fig.show='hold', fig.cap="Figure S2B: MA plot (Global normalization)", cache=FALSE}
plot_ma(tt) +
  ggtitle(label = "Global normalization") + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  theme(legend.position = "none", plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12))
```

## Figure S2C

SUPT4H1 effect by gene biotype (after spike-in normalization).

```{r}
# Spike-in normalization
fit_spikein <- fit_hek293(dge, normalize = "spike-in")
tt <- topTable(fit_spikein, coef = "SUPT4H1", n = Inf)
```

```{r fig.width=8.5, fig.height=6, , fig.show='hold', fig.cap="Figure S2C: SUPT4H1 effect by gene biotype (Spike-in normalization)", cache=FALSE}
plot_biotypes(tt)
```

## Figure S2D

SUPT4H1 effect on RNA polymerase components (after spike-in normalization).
The error bars show the mean and the standard error for each RNA polymerase complex.

```{r fig.height = 3.5, fig.width = 3.5, fig.show='hold', fig.cap="Figure S2D: Polymerase subunits (Spike-in normalization)", cache=FALSE}
pol_genes <- data.frame(
  hgnc_symbol = c("POLR1A", "POLR1B", "POLR1C", "POLR1C", "POLR1D", "POLR1D",
                  "POLR1E", "POLR2A", "POLR2B", "POLR2C",
                  "POLR2D", "POLR2E", "POLR2F", "POLR2G", "POLR2H",
                  "POLR2I", "POLR2I", "POLR2J", "POLR2J2", "POLR2J3",
                  "POLR2K", "POLR2L", "POLR2M", "POLR3A", "POLR3B",
                  "POLR3C", "POLR3D", "POLR3E", "POLR3F", "POLR3G",
                  "POLR3H", "POLR3K", "POLRMT"),
  polymerase = c("I", "I", "I", "III", "I", "III", "I", "II", "II", "II",
                 "II", "II", "II", "II", "II", "II", "I", "II",
                 "II", "II", "II", "II", "II", "III", "III",
                 "III", "III", "III", "III", "III", "III", "III",
                 "MT"),
  stringsAsFactors = FALSE
) %>%
  dplyr::select(hgnc_symbol, polymerase) %>%
  dplyr::inner_join(tt, by = "hgnc_symbol")

ggplot(pol_genes, aes(x = polymerase, y = logFC)) +
  geom_jitter(width = 0.1, shape = 21, size = 3,
              fill = "orange", color = "darkgrey") +
  stat_summary(fun.data = "mean_se") +
  scale_y_continuous(limits = c(NA, 0)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("RNA polymerase complex") +
  ylab("log2 fold change\n(spike-in normalized)") +
  ggtitle("SUPT4H1 RNAi effect") +
  theme(panel.grid = element_blank())
```

# HEK293 cell line models, fit separately for each stable cell line

## Control cell line

```{r fig.height = 4.5, fig.width = 5.5, fig.show='hold', fig.cap="Figure 2C: SUPT4H1 effect in Control cell line (spike-in normalization)", cache=FALSE}
# Spike-in normalization
fit_spikein <- fit_hek293_lines(dge, line = "Control", normalize = "spike-in")
tt <- topTable(fit_spikein, coef = "SUPT4H1", n = Inf)
plot_ma(tt) +
ggtitle("Control cell line")
```

## DARP1 cell line

```{r fig.height = 4.5, fig.width = 5.5, fig.show='hold', fig.cap="Figure 2C: SUPT4H1 effect in DARP1 cell line (spike-in normalization)", cache=FALSE}
# Spike-in normalization
fit_spikein <- fit_hek293_lines(dge, line = "DARP1", normalize = "spike-in")
tt <- topTable(fit_spikein, coef = "SUPT4H1", n = Inf)
plot_ma(tt) +
ggtitle("DARP1 cell line")
```

## DARP2 cell line

```{r fig.height = 4.5, fig.width = 5.5, fig.show='hold', fig.cap="Figure 2C: SUPT4H1 effect in DARP2 cell line (spike-in normalization)", cache=FALSE}
# Spike-in normalization
fit_spikein <- fit_hek293_lines(dge, line = "DARP2", normalize = "spike-in")
tt <- topTable(fit_spikein, coef = "SUPT4H1", n = Inf)
plot_ma(tt) +
ggtitle("DARP2 cell line")
```

## Figure S2F

Relationship between SUPT4H1 RNAi effect and annotated gene length in HEK293
cell lines (spike-in normalization).

```{r results='hide'}
fit <- fit_hek293(dge = dge, normalize = "spike-in")
gene_lengths <- read.csv(system.file("extdata", "gene_length.csv.gz", 
                                      package = "SUPT4H1"))
fit$genes$length <- gene_lengths[
  match(fit$genes$entrez_id, gene_lengths$entrez_id), "length"] / 1000
tt <- topTable(fit, coef = "SUPT4H1", n = Inf, sort.by = "none")
```

```{r fig.width=8.5, fig.height=6, fig.show='hold', fig.cap="Figure S5: SUPT4H1 effect is not gene-length dependent", cache=FALSE} 
par(mar = c(5, 6, 4, 2))
with(tt, plot(
  x = length,
  y = logFC,
  pch = 19,
  log = "x",
  col = adjustcolor("black", alpha.f = 0.25),
  ylim = c(-5, 2),
  xlim = c(0.5, 0.5e4),
  xlab = "Annotated gene length (kb)",
  ylab = "SUPT4H1 RNAi effect\n(log2 fold change)",
  xaxt = "n",
  cex.lab = 1.5,
  cex.axis = 1.5
))
ticks = 10^seq(0, 4, by = 1)
axis(1, at = ticks, labels = prettyNum(ticks), cex.axis = 1.5)

abline(h = 0, lty = 2)
with(subset(tt, hgnc_symbol == "SUPT4H1"), {
  points(length, logFC, col = "firebrick", pch = 19, cex = 1.5)
  text(length, logFC, col = "firebrick", labels = c("SUPT4H1"), pos = 4, 
       cex = 1.5)
})
loess_fit <- with(
  subset(tt, length > 1 & length < 1e3),
  lowess(length, logFC)
)
```

